---
title: |
  | Supplementary material: Dynamic modeling of corporate credit ratings and defaults
  | Analysis of synthetic data
author: "Laura Vana"
output: 
  pdf_document:
      toc: true
header-includes:
  - \usepackage{multicol}
  - \usepackage{float} 
  - \usepackage{tabularx}  
  - \usepackage{longtable} 

---
```{r helper}
logmeanexp <- function (x) {
  mx <- max(x)
  m <- mx+log(mean(exp(x-mx)))
  m
} 
```

# Data set

The synthetic data set `synthetic_dat.rda` has been generated using 
the **synthpop** R package (Nowok et. al, 2016) by replacing all observations
with values simulated from probability distributions specified to 
preserve key features of the actual observed data.

The data set contains 19952 rows and 13 columns:

* `firm_id`: integer containing the firm id, maximum is 2528

* `year_id`: integer containing the year id, maximum is 19

* `R1`: integer corresponding to the rating classes assigned by rater 1

* `R2`: integer corresponding to the rating classes assigned by rater 2

* `R3`: integer corresponding to the rating classes assigned by rater 3

* `D` : binary indicator for default

* `X1` to `X7`: standardized covariates.

# Estimating the models -- full sample and out-of-time 

```{r}
nchains <- 5
niter <- 2000L

########################################################
########################################################
### CODE FOR ESTIMATION WITH rstan ON SYNTHETIC DATA ###
########################################################
########################################################
library(rstan)
library(coda)

## ----load_data-------------------------------------------------------------------
load("synthetic_dat.rda")
dat <- synthetic_dat

take.subsample <- TRUE

if (take.subsample) {
  ## take only p% of the firms 
  p <- 0.1
  set.seed(123)
  nfirms <- length(unique(dat$firm_id))
  firms_def_1 <- unique(dat$firm_id[dat$D == 1])
  firms_def_0 <- unique(dat$firm_id[dat$D == 0])
  firm_ids_small <- c(sample(firms_def_1, p * length(firms_def_1)),
                      sample(firms_def_0, p * length(firms_def_0)))
  dat <- subset(dat, firm_id %in% firm_ids_small)
}

## ---------------- stan_models ----------------------
models <- list(
  list(## Model S1
    model = "../Stan/Model_S1_logit_priors-eps-N_bias-0.stan",
    pars  = c("beta", "beta0",
              "theta1", "theta2", "theta3",
              "psi", 
              "log_lik_all", 
              "log_lik_D_test", "log_lik_R1_test", "log_lik_R2_test", "log_lik_R3_test")),
  list(## Model S2
    model = "../Stan/Model_S2_logit_priors-eps-N_bias-diffgammas.stan",
    pars  = c("beta", "gamma1", "gamma2", "gamma3",  "beta0", 
              "theta1", "theta2", "theta3",
              "psi",
              "log_lik_all", 
              "log_lik_D_test", "log_lik_R1_test", "log_lik_R2_test", "log_lik_R3_test")),
  list(## Model D1
    model = "../Stan/Model_D1_logit_priors-a-HN-b-AR1-eps-AR1_bias-0.stan",
    pars  = c("beta", "beta0", 
              "theta1", "theta2", "theta3",
              "q2",
              "b", "omega", "phib",
              "rho", "psi",
              "log_lik_all", 
              "log_lik_D_test", "log_lik_R1_test", "log_lik_R2_test", "log_lik_R3_test"
    )),
  list(## Model D2
    model = "../Stan/Model_D2_logit_priors-a-HN-b-AR1-eps-AR1_bias-diffgammas.stan",
    pars  = c("beta", "gamma1", "gamma2", "gamma3",  "beta0",
              "theta1", "theta2", "theta3",
              "q2", 
              "b", "omega", "phib", 
              "rho", "psi",
              "log_lik_all", 
              "log_lik_D_test", "log_lik_R1_test", "log_lik_R2_test", "log_lik_R3_test"
    )),
  list(## Model PM
    model = "../Stan/Model_PM_logit_priors-a-HN-b-AR1-eps-AR1_bias-diffbetas-delta-AR1.stan",
    pars  = c("beta", "gamma1", "gamma2", "gamma3", "beta0",
              "theta1", "theta2", "theta3",
              "q2", 
              "b", "omega", "phib",
              "rho", "psi",
              "delta", "lambda", "phidelta",
              "log_lik_all", 
              "log_lik_D_test", "log_lik_R1_test", "log_lik_R2_test", "log_lik_R3_test")))


##############################
### OUT OF SAMPLE ANALYSIS ###
##############################
# for the one step ahead out of sample exercise, we train the model on the
# data containing years 1 : t and check the log predictive 
# likelihoods for the following year t + 1

end_of_training_samples <- max(dat$year_id)
number_of_models        <- length(models)
for (t in 13:end_of_training_samples) {
  for (s in seq_len(number_of_models)){
    print(c(t, s))
    # select only observations in years 1:(t + 1)
    dat_tmp <- subset(dat, year_id %in% seq_len(t + 1))
    dat_tmp$firm_id <- as.numeric(factor(dat_tmp$firm_id))
    
    id_train <- which(dat_tmp$year_id <= t)
    id_test  <- which(dat_tmp$year_id ==  t + 1)
    
    N1 <- length(id_train)
    N2 <- length(id_test)
    
    # matrix of ratings in training sample
    y_train <- dat_tmp[id_train, c("R1", "R2", "R3")]
    # matrix of ratings in test sample
    y_test  <- dat_tmp[id_test,  c("R1", "R2", "R3")]
    
    begin_eps <- seq_len(N1 + N2)[!duplicated(dat_tmp$firm_id)]
    end_eps   <- c(begin_eps[-1] - 1, N1 + N2)
    
    DATA <- list(
      N             = N1 + N2, 
      N1            = N1, 
      N2            = N2,
      NR1_train     = sum(!is.na(y_train[, 1])),
      NR2_train     = sum(!is.na(y_train[, 2])),
      NR3_train     = sum(!is.na(y_train[, 3])),
      NR1_test      = sum(!is.na(y_test[, 1])),
      NR2_test      = sum(!is.na(y_test[, 2])),
      NR3_test      = sum(!is.na(y_test[, 3])),
      P             = 7,  
      J             = 3,
      I             = length(unique(dat_tmp$firm_id)),
      T             = max(dat_tmp$year_id),
      id_test       = N1 + seq_len(N2),
      begin_eps     = begin_eps,
      end_eps       = end_eps,
      firm_id       = dat_tmp$firm_id,
      firm_id_train = dat_tmp$firm_id[id_train],
      firm_id_test  = dat_tmp$firm_id[id_test],
      year_id       = dat_tmp$year_id,
      year_id_train = dat_tmp$year_id[id_train],
      year_id_test  = dat_tmp$year_id[id_test],
      id_obs_R1_train = which(!is.na(y_train[, 1])),
      id_obs_R2_train = which(!is.na(y_train[, 2])),
      id_obs_R3_train = which(!is.na(y_train[, 3])),
      id_obs_R1_test  = which(!is.na(y_test[, 1])),
      id_obs_R2_test  = which(!is.na(y_test[, 2])),
      id_obs_R3_test  = which(!is.na(y_test[, 3])),
      K1 = max(y_train[, 1], na.rm = TRUE),
      K2 = max(y_train[, 2], na.rm = TRUE),
      K3 = max(y_train[, 3], na.rm = TRUE),
      y1_train = y_train[!is.na(y_train[, 1]), 1],
      y2_train = y_train[!is.na(y_train[, 2]), 2],
      y3_train = y_train[!is.na(y_train[, 3]), 3],
      y1_test = y_test[!is.na(y_test[, 1]), 1],
      y2_test = y_test[!is.na(y_test[, 2]), 2],
      y3_test = y_test[!is.na(y_test[, 3]), 3],
      D_train = dat_tmp$D[id_train], 
      D_test  = dat_tmp$D[id_test],
      x_train = dat_tmp[id_train, grep("X", colnames(dat_tmp))], 
      x_test  = dat_tmp[id_test , grep("X", colnames(dat_tmp))]
    )
    # ------------- prepare files and folders ---------------       
    model_name <- gsub("\\..*", "",  gsub("../Stan/", "", models[[s]][["model"]]))
    res_folder <- "Results_Synthetic_Data"
    pars <-  models[[s]][["pars"]]
    if (!dir.exists(res_folder)) dir.create(res_folder, recursive = TRUE)
    if (t == max(dat$year_id)) {
      file_out <- paste0(res_folder, "/", sprintf("results_%s_fullsample.rda", model_name))
      pars <- gsub("_test", "", pars)
    } else {
      file_out <- paste0(res_folder, "/", sprintf("OOS_results_%s_testperiod_%i.rda", model_name, t + 1))
    }
    # ------------- analysis -------------
    fit <- stan(file = models[[s]][["model"]],
                pars = models[[s]][["pars"]],
                data = DATA, 
                chains = nchains, 
                iter = niter,
                save_warmup = FALSE, 
                control=list(adapt_delta = 0.99))
    
    save(fit, file = file_out)
  }
}

```

## Computing the LOO values


```{r}
models <- c(
  "Model_S1_logit_priors-eps-N_bias-0",
  "Model_S2_logit_priors-eps-N_bias-diffgammas",
  "Model_D1_logit_priors-a-HN-b-AR1-eps-AR1_bias-0",
  "Model_D2_logit_priors-a-HN-b-AR1-eps-AR1_bias-diffgammas",
  "Model_PM_logit_priors-a-HN-b-AR1-eps-AR1_bias-diffbetas-delta-AR1"
)

loo_list <- list()
i <- 0
for (m in models) {
  i <- i + 1
  load(sprintf("Results_Synthetic_Data/results_%s_fullsample.rda", m))
  ll <- loo::extract_log_lik(fit, parameter_name = "log_lik_all")
  loo_list[[i]] <- loo::loo(ll, k_threshold=0.7)
}
loo_syn_data <- loo::loo_compare(loo_list)
loo_syn_data
```

## Out-of-time measures

We perform the one step ahead out-of time analysis. We compute the joint log predictive densities, as well as the measures based on the conditional PDs.

```{r}
#---------------------
test_period <- 14:19



## Compute log predictive density 
lppd <- lapply(models, function(x) {
  sapply(seq_along(test_period), function(ti){
    FILE <- sprintf("Results_Synthetic_Data/OOS_results_%s_testperiod_%i.rda", 
                    x, test_period[ti])
    print(FILE)
    load(FILE)
    dat_test <- dat[dat$year_id %in% test_period[ti], ] 
    log_lik_D <- extract(fit, pars = "log_lik_D_test")[[1]]
    
    log_lik_S_with_NA <- log_lik_M_with_NA <- log_lik_F_with_NA <-
      matrix(0, ncol = ncol(log_lik_D), nrow = nrow(log_lik_D))
    
    log_lik_S <- extract(fit, pars = "log_lik_R1_test")[[1]]
    log_lik_M <- extract(fit, pars = "log_lik_R2_test")[[1]]
    log_lik_F <- extract(fit, pars = "log_lik_R3_test")[[1]]
    
    log_lik_S_with_NA[, !is.na(dat_test$SPR)] <- log_lik_S
    log_lik_M_with_NA[, !is.na(dat_test$Moodys)] <- log_lik_M
    log_lik_F_with_NA[, !is.na(dat_test$Fitch)] <- log_lik_F
    
    log_lik_all <- log_lik_D + log_lik_S_with_NA + log_lik_M_with_NA + log_lik_F_with_NA
    
    log_pred_all <- mean(log(colMeans(exp(log_lik_all))))

    log_pred_D   <- mean(log(colMeans(exp(log_lik_D))))
    
    log_pred_S   <- mean(log(colMeans(exp(log_lik_S))))
    log_pred_M   <- mean(log(colMeans(exp(log_lik_M))))
    log_pred_F   <- mean(log(colMeans(exp(log_lik_F))))
    res <- c(log_pred_all,
             log_pred_D, log_pred_S,
             log_pred_M, log_pred_F)
    return(res)
  })
})

names(lppd) <- models
save(lppd, file = "Results_Synthetic_Data/lppd_1ahead.rds")

conditional_probabilities <- lapply(models, function(x) {
  sapply(seq_along(test_period), function(ti){
    FILE <- sprintf("Results_Synthetic_Data/OOS_results_%s_testperiod_%i.rda", 
                    x, test_period[ti])
    print(FILE)
    load(FILE)
    dat_test <- dat[dat$year_id %in% test_period[ti], ] 
    
    log_lik_D <- extract(fit, pars = "log_lik_D_test")[[1]]
    
    log_lik_S_with_NA <- log_lik_M_with_NA <- log_lik_F_with_NA <-
      matrix(0, ncol = ncol(log_lik_D), nrow = nrow(log_lik_D))
    
    log_lik_S <- extract(fit, pars = "log_lik_R1_test")[[1]]
    log_lik_M <- extract(fit, pars = "log_lik_R2_test")[[1]]
    log_lik_F <- extract(fit, pars = "log_lik_R3_test")[[1]]
    
    log_lik_S_with_NA[, !is.na(dat_test$SPR)] <- log_lik_S
    log_lik_M_with_NA[, !is.na(dat_test$Moodys)] <- log_lik_M
    log_lik_F_with_NA[, !is.na(dat_test$Fitch)] <- log_lik_F
    
    log_lik_rat <-  log_lik_S_with_NA + log_lik_M_with_NA + log_lik_F_with_NA
    
    
    D_tmp <- dat_test$D
    # ---------
    #  Compute P(D=1, S, M, F)
    # ---------
    # Compute log  P(D=1)
    ## P(D=1|zeta)
    probs_D <- exp(log_lik_D)
    probs_D[, D_tmp == 0] <- 1 - probs_D[, D_tmp == 0]
    probs_D[probs_D == 0] <- .Machine$double.eps
    probs_D[probs_D == 1] <- 1 - .Machine$double.eps
    log_lik_D_1 <- log(probs_D)
    
    log_prob_all <- (log_lik_D_1 + log_lik_rat)
    # ---------
    #  Compute P(S, M, F)
    # ---------
    apply(log_prob_all, 2, logmeanexp)
    prob_rat <- exp(log_lik_rat)
    log_pd_conditional <- apply(log_prob_all, 2, logmeanexp) - apply(log_lik_rat, 2, logmeanexp) 
    return(log_pd_conditional)
  })
})
    



res_BR <- sapply(conditional_probabilities, 
                  function(y) {
                    tab <- sapply(seq_along(test_period), function(ti){ 
                      lp <- y[[ti]]
                      D_test <-  dat[dat$year_id %in% test_period[ti], "D"] 
                      mean((exp(lp) - D_test)^2) 
                    })
                    tab
                  })



## Area under precision-recall curve
if (!require("PRROC")) install.packages("PRROC"); library("PRROC")
res_AUC_PRAR <- sapply(conditional_probabilities, 
                 function(y) {
                   tab <- sapply(seq_along(test_period), function(ti){ 
                     lp <- y[[ti]]
                     D_test <- dat[dat$year_id %in% test_period[ti], "D"] 
                     pr.curve(scores.class0 = lp[D_test==1], scores.class1 = lp[D_test==0],
                              curve = FALSE)$auc.integral
                     
                   })
                   tab
                 })


colnames(res_BR) <- colnames(res_AUC_PRAR) <-  models

save(res_BR, res_AUC_PRAR, file = "Results_Synthetic_Data/pd_cond_measures_1ahead.rda")
```

```{r}
id_years <- seq_along(test_period)
lppd <- lapply(lppd, function(x) x[c(1),,drop=FALSE ])
tab_loo <- t(loo_syn_data[,1:2])[,paste0("model", 1:5)]
tab <- rbind(sapply(lppd, function(x) rowMeans(x[, id_years,drop=F])),
             apply(sqrt(res_BR), 2, function(x) mean(x[id_years])),
              apply(res_AUC_PRAR, 2, function(x) mean(x[id_years], na.rm = TRUE)))
tab
```



#  References

